noam_final_analyses

### 2024-09-10 noam_final_analyses.Rmd

```{r beta dispersion}
full_site_species_matrix <- as.matrix(site_full[ , -which(names(site_full) %in% meta_cols)])
full_site_species_matrix <- full_site_species_matrix/site_full$hard_cover # convert from percent to proportion of hard cover
rownames(full_site_species_matrix) <- paste(site_full$year,"-", site_full$site_habitat, sep = "")

@@

full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% colnames(site_meta_df))])
full_site_macro_species_matrix <- full_site_macro_species_matrix/site_macro$hard_cover # convert from percent to proportion of hard cover
rownames(full_site_macro_species_matrix) <- paste(site_macro$year,"-", site_macro$site_habitat, sep = "")

full_site_macro_species_matrix <- asin(sqrt(full_site_macro_species_matrix)) # arcsin squareroot transform

full_site_bray <- vegdist(full_site_species_matrix, "bray") 
full_site_macro_bray <- vegdist(full_site_macro_species_matrix, "bray")

full_site_betadisper <- vegan::betadisper(full_site_bray, site_full$hab_year)
site_full$betadispersion <- full_site_betadisper$distances[match(site_full$stratum, names(full_site_betadisper$distances))]
site_full$betadispersion_trans <- sv_trans(site_full$betadispersion, site_full)

@@

full_site_macro_betadisper <- vegan::betadisper(full_site_macro_bray, site_macro$hab_year)
full_subset_site_macro$betadispersion <- full_site_macro_betadisper$distances[match(full_subset_site_macro$stratum, names(full_site_macro_betadisper$distances))]
full_subset_site_macro$betadispersion_trans <- sv_trans(full_subset_site_macro$betadispersion, full_subset_site_macro)

@@

@@

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_site_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

@@

alpha_diversity_quad <- full_join(alpha_diversity_quad, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_quad, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

@@

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_quad_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

@@

@@

site_macro_shannon_mod <- glmmTMB(log(shannon +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed
summary(site_macro_shannon_mod)
Anova(site_macro_shannon_mod)
#                Chisq Df Pr(>Chisq)    
# year          0.0133  1   0.908348    
# habitat      97.6274  3  < 2.2e-16 ***
# year:habitat 14.0548  3   0.002831 ** 
hist(residuals(site_macro_shannon_mod)) # looks good
plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod)) # generally fine

@@

@@

## Backreef - includes all 
# backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
# 
# backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + # offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has # convergence issues
# backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
# model.sel(backreef_all_richness_dredge) # top model includes year and coral 
# backreef_all_richness_top_mod <- glmmTMB(richness ~ year  + offset(log(hard_cover)) + (1|site), family = poisson, data = # backreef_all_site, na.action = na.pass)
# summary(backreef_all_richness_top_mod)
# hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
# plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine

@@

>>>>

```{r beta dispersion}
site_full$stratum <- paste(site_full$year,"-", site_full$site_habitat, sep = "") # for merge
site_macro$stratum <- paste(site_macro$year,"-", site_macro$site_habitat, sep = "") # for merge

full_site_species_matrix <- as.matrix(site_full[ , -which(names(site_full) %in% c(meta_cols, "stratum" ))])
full_site_species_matrix <- full_site_species_matrix/site_full$hard_cover # convert from percent to proportion of hard cover
rownames(full_site_species_matrix) <- paste(site_full$year,"-", site_full$site_habitat, sep = "")

@@

full_site_macro_species_matrix <- as.matrix(site_macro[ , -which(names(site_macro) %in% c(meta_cols, "stratum" ))])
full_site_macro_species_matrix <- full_site_macro_species_matrix/site_macro$hard_cover # convert from percent to proportion of hard cover
rownames(full_site_macro_species_matrix) <- paste(site_macro$year,"-", site_macro$site_habitat, sep = "")

full_site_macro_species_matrix <- asin(sqrt(full_site_macro_species_matrix)) # arcsin squareroot transform
# some rows are all 0s
full_site_macro_rows_to_remove <- rowSums(full_site_macro_species_matrix) == 0
full_site_macro_species_matrix <- full_site_macro_species_matrix[!full_site_macro_rows_to_remove,] 
full_site_bray <- vegdist(full_site_species_matrix, "bray") 
full_site_macro_bray <- vegdist(full_site_macro_species_matrix, "bray")

full_site_betadisper <- vegan::betadisper(full_site_bray, site_full$hab_year)
site_full$betadispersion <- full_site_betadisper$distances[match(site_full$stratum, names(full_site_betadisper$distances))]
site_full$betadispersion_trans <- sv_trans(site_full$betadispersion)

@@

full_subset_site_macro <- site_macro[site_macro$stratum %in% rownames(full_site_macro_species_matrix),]
full_site_macro_betadisper <- vegan::betadisper(full_site_macro_bray, site_macro_subset$hab_year)
full_subset_site_macro$betadispersion <- full_site_macro_betadisper$distances[match(full_subset_site_macro$stratum, names(full_site_macro_betadisper$distances))]
full_subset_site_macro$betadispersion_trans <- sv_trans(full_subset_site_macro$betadispersion)

@@

# add beta dispersion
alpha_diversity_site <- full_join(alpha_diversity_site, site_full[,c("betadispersion_trans", "site_hab_year")], by = "site_hab_year")
alpha_diversity_site$year_factor <- as.factor(as.character(alpha_diversity_site$year))

@@

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, full_subset_site_macro[,c("betadispersion_trans", "site_hab_year")], by = "site_hab_year")
alpha_diversity_site_macro$year_factor <- as.factor(as.character(alpha_diversity_site_macro$year))

@@

alpha_diversity_quad <- full_join(alpha_diversity_quad, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
alpha_diversity_quad$year_factor <- as.factor(as.character(alpha_diversity_quad$year))

@@

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
alpha_diversity_quad_macro$year_factor <- as.factor(as.character(alpha_diversity_quad_macro$year))

```{r all algae cover mods}
site_full_cover_mod <- glmmTMB(sv_trans( all_algae_cover_hard/100) ~ year*habitat + (1|site) , family = beta_family(),
                             data = alpha_diversity_site) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_full_cover_mod) # decent variance in random effect
Anova(site_full_cover_mod) 
#                Chisq Df Pr(>Chisq)    
# year          0.9983  1     0.3177    
# habitat      43.0233  3  2.433e-09 ***
# year:habitat 49.8146  3  8.750e-11 ***
# habitat more important than time when time is numeric
# check_mod
hist(residuals(site_full_cover_mod)) # negative skew

plot(residuals(site_full_cover_mod) ~ fitted(site_full_cover_mod)) # negative trend

emtrends(site_full_cover_mod, pairwise ~  habitat, var = "year")
# backreef not different from fringing, outer10 not different from outer17
site_full_cover_emm <- emmip(site_full_cover_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_cover_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = all_algae_cover_hard/100, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_cover_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_full_cover_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "Cover (prop)") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_full_cover_mod_factor <- glmmTMB(sv_trans( all_algae_cover_hard/100) ~ year_factor*habitat + 
                                        (1|site) + ar1(year_factor + 0 | site),
                                  family = beta_family(), data = alpha_diversity_site)

summary(site_full_cover_mod_factor) # low variance in random effect but strong temporal autocorrelation
Anova(site_full_cover_mod_factor) 
#                 Chisq Df Pr(>Chisq)    
# year_factor          93.081 16  6.745e-13 ***
# habitat             113.174  3  < 2.2e-16 ***
# year_factor:habitat  63.853 46    0.04171 *  
hist(residuals(site_full_cover_mod_factor)) # still neg skew but not as bad

plot(residuals(site_full_cover_mod_factor) ~ fitted(site_full_cover_mod_factor)) # strong patterning

site_full_cover_factor_emm <- plot(emmeans(site_full_cover_mod_factor, ~  habitat*year_factor), plotit = F) 
site_full_cover_factor_emm$year <- as.numeric(levels(site_full_cover_factor_emm$year_factor)) # for plotting

(site_factor_cover_plot <- 
  ggplot() +
  geom_point(data = site_full_cover_factor_emm,
             aes(x = year, y = inv_logit(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
  geom_line(data = site_full_cover_factor_emm, aes(x = year, y = inv_logit(the.emmean), colour = habitat, group = habitat),
            position = position_dodge(0.4)) +
  geom_errorbar(data = site_full_cover_factor_emm,
                aes(x = year, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), colour = habitat), width = 0.2,
                position = position_dodge(0.4)) +
  geom_segment(aes(y = 1, yend = 1, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "cover") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```{r macroalgaealgae cover mods}
site_macrofull_cover_mod <- glmmTMB(sv_trans( macroalgae_cover_hard/100) ~ year*habitat + (1|site) , family = beta_family(),
                             data = alpha_diversity_site_macro) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_macrofull_cover_mod) # less variance in random effect than before
Anova(site_macrofull_cover_mod) 
#                Chisq Df Pr(>Chisq)    
# year          3.9366  1   0.047247 *  
# habitat      32.7335  3  3.666e-07 ***
# year:habitat 13.0316  3   0.004569 ** 
hist(residuals(site_macrofull_cover_mod)) # positive skew

plot(residuals(site_macrofull_cover_mod) ~ fitted(site_macrofull_cover_mod)) # trend likely due to beta reg

emtrends(site_macrofull_cover_mod, pairwise ~  habitat, var = "year")
# backreef different from outer10
site_macrofull_cover_emm <- emmip(site_macrofull_cover_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_macrobase_cover_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro,
             aes(x = year, y = macroalgae_cover_hard/100, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macrofull_cover_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_macrofull_cover_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 0.7, yend = 0.7, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "Cover (prop)") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_macrofull_cover_mod_factor <- glmmTMB(sv_trans( macroalgae_cover_hard/100) ~ year_factor*habitat + 
                                        (1|site) + ar1(year_factor + 0 | site),
                                  family = beta_family(), data = alpha_diversity_site_macro)

summary(site_macrofull_cover_mod_factor) # low variance in random effect but strong temporal autocorrelation
Anova(site_macrofull_cover_mod_factor) 
#                 Chisq Df Pr(>Chisq)    
# year_factor         35.931 16   0.002958 ** 
# habitat             36.929  3  4.764e-08 ***
# year_factor:habitat 98.623 46  1.056e-05 ***
hist(residuals(site_macrofull_cover_mod_factor)) # still pos skew but not as bad

plot(residuals(site_macrofull_cover_mod_factor) ~ fitted(site_macrofull_cover_mod_factor)) # similar to numeric

site_macrofull_cover_factor_emm <- plot(emmeans(site_macrofull_cover_mod_factor, ~  habitat*year_factor), plotit = F) 
site_macrofull_cover_factor_emm$year <- as.numeric(levels(site_macrofull_cover_factor_emm$year_factor)) # for plotting

(site_macrofactor_cover_plot <- 
  ggplot() +
  geom_point(data = site_macrofull_cover_factor_emm,
             aes(x = year, y = inv_logit(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
  geom_line(data = site_macrofull_cover_factor_emm, aes(x = year, y = inv_logit(the.emmean), colour = habitat, group = habitat),
            position = position_dodge(0.4)) +
  geom_errorbar(data = site_macrofull_cover_factor_emm,
                aes(x = year, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), colour = habitat), width = 0.2,
                position = position_dodge(0.4)) +
  geom_segment(aes(y = 0.5, yend = 0.5, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
    labs(x = "Year", y = "cover") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

@@

```{r all algae alpha diversity site level year factor}

site_full_richness_mod_factor <- glmmTMB(richness ~ year_factor*habitat + (1|site) + ar1(year_factor + 0 | site) + offset(log(hard_cover)),
                                  family = poisson, data = alpha_diversity_site)

summary(site_full_richness_mod_factor) # low variance in random effect but strong temporal autocorrelation
Anova(site_full_richness_mod_factor) 
#                 Chisq Df Pr(>Chisq)    
# year_factor          93.081 16  6.745e-13 ***
# habitat             113.174  3  < 2.2e-16 ***
# year_factor:habitat  63.853 46    0.04171 *  
hist(residuals(site_full_richness_mod_factor)) # looks great

# striping/patterining expected because of discrete response var.
plot(residuals(site_full_richness_mod_factor) ~ fitted(site_full_richness_mod_factor))

site_full_richness_factor_emm <- plot(emmeans(site_full_richness_mod_factor, ~  habitat*year_factor), plotit = F) 
site_full_richness_factor_emm$year <- as.numeric(levels(site_full_richness_factor_emm$year_factor)) # for plotting

(site_factor_richness_plot <- 
  ggplot() +
  geom_point(data = site_full_richness_factor_emm,
             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
  geom_line(data = site_full_richness_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
            position = position_dodge(0.4)) +
  geom_errorbar(data = site_full_richness_factor_emm,
                aes(x = year, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL), colour = habitat), width = 0.2,
                position = position_dodge(0.4)) +
  geom_segment(aes(y = 22, yend = 22, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_full_shannon_mod_factor <- glmmTMB(log(shannon+1) ~ year_factor*habitat + (1|site) + ar1(year_factor + 0 | site),
                                        data = alpha_diversity_site) # model convergence - goes away if you drop site random effect
# results are identical between including RE and not

summary(site_full_shannon_mod_factor) # low variance in random effect but strong temporal autocorrelation
Anova(site_full_shannon_mod_factor) 
#                 Chisq Df Pr(>Chisq)    
# year_factor         129.09 16  < 2.2e-16 ***
# habitat             156.55  3  < 2.2e-16 ***
# year_factor:habitat 152.84 46  2.162e-13 ***
hist(residuals(site_full_shannon_mod_factor)) # a few points drive slight negative skew but good on the whole
plot(residuals(site_full_shannon_mod_factor) ~ fitted(site_full_shannon_mod_factor))# looks good

site_full_shannon_factor_emm <- plot(emmeans(site_full_shannon_mod_factor, ~  habitat*year_factor), plotit = F) 
site_full_shannon_factor_emm$year <- as.numeric(levels(site_full_shannon_factor_emm$year_factor)) # for plotting

(site_factor_shannon_plot <- 
  ggplot() +
  geom_point(data = site_full_shannon_factor_emm,
             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
  geom_line(data = site_full_shannon_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
            position = position_dodge(0.4)) +
  geom_errorbar(data = site_full_shannon_factor_emm,
                aes(x = year, ymin = exp(lower.CL), ymax = exp(upper.CL), colour = habitat), width = 0.2,
                position = position_dodge(0.4)) +
  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

@@

site_macro_shannon_mod <- glmmTMB(log(shannon +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site_macro) 
# sites with shannon div = 0 so log+1 transformed
summary(site_macro_shannon_mod)
Anova(site_macro_shannon_mod)
#                Chisq Df Pr(>Chisq)    
# year           6.7643  1    0.00930 ** 
# habitat      104.2995  3    < 2e-16 ***
# year:habitat   9.1657  3    0.02717 *  
qqnorm(residuals(site_macro_shannon_mod)) # slight negative skew
plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod)) # generally fine

@@

```{r macroalgae alpha diversity site level year factor}

site_macrofull_richness_mod_factor <- glmmTMB(richness ~ year_factor*habitat + (1|site) + ar1(year_factor + 0 | site) +
                                              offset(log(hard_cover)), family = poisson, data = alpha_diversity_site_macro)

summary(site_macrofull_richness_mod_factor) # slightly lower temporal autocorrelation than before but still high
Anova(site_macrofull_richness_mod_factor) 
#                 Chisq Df Pr(>Chisq)    
# year_factor         80.475 16  1.366e-10 ***
# habitat             92.482  3  < 2.2e-16 ***
# year_factor:habitat 54.467 46     0.1835    
qqnorm(residuals(site_macrofull_richness_mod_factor)) # some positive skew

# striping/patterining expected because of discrete response var.
plot(residuals(site_macrofull_richness_mod_factor) ~ fitted(site_macrofull_richness_mod_factor))

site_macrofull_richness_factor_emm <- plot(emmeans(site_macrofull_richness_mod_factor, ~  habitat*year_factor), plotit = F) 
site_macrofull_richness_factor_emm$year <- as.numeric(levels(site_macrofull_richness_factor_emm$year_factor)) # for plotting

(site_macrofactor_richness_plot <- 
  ggplot() +
  geom_point(data = site_macrofull_richness_factor_emm,
             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
  geom_line(data = site_macrofull_richness_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
            position = position_dodge(0.4)) +
  geom_errorbar(data = site_macrofull_richness_factor_emm,
                aes(x = year, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL), colour = habitat), width = 0.2,
                position = position_dodge(0.4)) +
  geom_segment(aes(y = 22, yend = 22, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_macrofull_shannon_mod_factor <- glmmTMB(log(shannon+1) ~ year_factor*habitat + (1|site) + ar1(year_factor + 0 | site),
                                        data = alpha_diversity_site_macro) # no convergence issues

summary(site_macrofull_shannon_mod_factor) # fairly low temporal ac
Anova(site_macrofull_shannon_mod_factor) 
#                 Chisq Df Pr(>Chisq)    
# year_factor          88.821 16  4.122e-12 ***
# habitat             144.112  3  < 2.2e-16 ***
# year_factor:habitat  58.923 46    0.09568 .  
qqnorm(residuals(site_macrofull_shannon_mod_factor)) # a few points wonky but pretty good
plot(residuals(site_macrofull_shannon_mod_factor) ~ fitted(site_macrofull_shannon_mod_factor))# looks good

site_macrofull_shannon_factor_emm <- plot(emmeans(site_macrofull_shannon_mod_factor, ~  habitat*year_factor), plotit = F) 
site_macrofull_shannon_factor_emm$year <- as.numeric(levels(site_macrofull_shannon_factor_emm$year_factor)) # for plotting

(site_macrofactor_shannon_plot <- 
  ggplot() +
  geom_point(data = site_macrofull_shannon_factor_emm,
             aes(x = year, y = exp(the.emmean), colour = habitat), size = 2, position = position_dodge(0.4)) +
  geom_line(data = site_macrofull_shannon_factor_emm, aes(x = year, y = exp(the.emmean), colour = habitat, group = habitat),
            position = position_dodge(0.4)) +
  geom_errorbar(data = site_macrofull_shannon_factor_emm,
                aes(x = year, ymin = exp(lower.CL), ymax = exp(upper.CL), colour = habitat), width = 0.2,
                position = position_dodge(0.4)) +
  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

@@

## Backreef - includes all 
backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors
# 
# backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + # offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has # convergence issues
# backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
# model.sel(backreef_all_richness_dredge) # top model includes year and coral 
# backreef_all_richness_top_mod <- glmmTMB(richness ~ year  + offset(log(hard_cover)) + (1|site), family = poisson, data = # backreef_all_site, na.action = na.pass)
# summary(backreef_all_richness_top_mod)
# hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
# plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine

@@

```{r stepwise models beta dispersion site level, fig.height = 8, fig.width=12}

## FRINGING - exclude wave power

fringing_all_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass, family = beta_family())
fringing_all_betadispersion_dredge <- dredge(fringing_all_betadispersion_mod_universal) 
model.sel(fringing_all_betadispersion_dredge) # top model is just intercept, next best is just year
fringing_all_betadispersion_top_mod <- glmmTMB(betadispersion_trans ~ year + (1|site), data = fringing_all_site, na.action = na.pass, family = beta_family())

qqnorm(residuals(fringing_all_betadispersion_top_mod))# good

plot(residuals(fringing_all_betadispersion_top_mod) ~ fitted(fringing_all_betadispersion_top_mod)) # good
summary(fringing_all_betadispersion_top_mod) # not even sig

## Backreef - includes all

backreef_all_betadispersion_mod_universal <- glmmTMB(betadispersion_trans ~ year + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass, family = beta_family())
backreef_all_betadispersion_dredge <- dredge(backreef_all_betadispersion_mod_universal) 
model.sel(backreef_all_betadispersion_dredge) #  best includes just nitrogen
backreef_all_betadispersion_top_mod <- glmmTMB(betadispersion_trans ~ mean_n  + (1|site), data = backreef_all_site, na.action = na.pass, family = beta_family())
qqnorm(residuals(backreef_all_betadispersion_top_mod)) #  a little wonky on the tail but not bad
plot(residuals(backreef_all_betadispersion_top_mod) ~ fitted(backreef_all_betadispersion_top_mod)) # looks fine
summary(backreef_all_betadispersion_top_mod) # borderline
Anova(backreef_all_betadispersion_top_mod) # X2 = 3.5, p = 0.06

backreef_emm <-as.data.frame(emmeans(backreef_all_betadispersion_top_mod, ~mean_n,
                                     at = list(mean_n = seq(min(backreef_all_site$mean_n),
                                                            max(backreef_all_site$mean_n),
                                                            0.01))))

ggplot()+
  geom_point(data = backreef_all_site, aes(x = mean_n, y = betadispersion_trans))+
  geom_line(data = backreef_emm, aes(x = mean_n, y = inv_logit(emmean))) +
  geom_ribbon(data = backreef_emm, aes(x = mean_n, ymin = inv_logit(asymp.LCL), ymax = inv_logit(asymp.UCL), alpha = 0.2), show.legend = F) +
  labs(x = "Mean N", y = "Backreef beta dispersion") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())
